Concurrent Programming with Harmony

### 2020-08-11 book/paper.tex
Latest

So far we have concentrated on critical sections to synchronize multiple
processes.  Certainly, preventing multiple processes from accessing
certain code at the same time simplifies how to think about synchronization.
However, it can lead to starvation.  Even in the absence of starvation,
if some process is slow for some reason while being in the critical section,
the other processes have to wait for it to finish executing the critical section.
In this chapter, we will have a look at how one can develop concurrent
code in which processes do not have to wait for other processes.

@@

There are also various algorithms where read-only access is wait-free
but updates do require a lock---these are useful in situations where
most accesses are read-only and the performance of updates is less
critical~\cite{}.
\autoref{fig:lst1} and \autoref{fig:lst2} implements an ordered linked
list of integers without duplicates.
There are two update operations:
values can be added using \texttt{lst\_insert} or deleted using
\texttt{lst\_remove}.
There is also a read-only operation: \texttt{lst\_contains} checks if
a particular value is in the list.
\glsadd{linearizable}%

@@

\begin{figure}
\begin{code}
\harmonysource{lst1}
\end{code}
\caption{\harmonylink{code/lst1.hny} List with non-blocking read operations, part 1.}
\label{fig:lst1}
\end{figure}

\begin{figure}
\begin{code}
\harmonysource{lst2}
\end{code}
\caption{\harmonylink{code/lst2.hny} List with non-blocking read operations, part 2.}
\label{fig:lst2}
\end{figure}

>>>>

So far we have concentrated on critical sections to synchronize multiple
processes.  Certainly, preventing multiple processes from accessing
certain code at the same time simplifies how to think about synchronization.
However, it can lead to starvation.  Even in the absence of starvation,
if some process is slow for some reason while being in the critical section,
the other processes have to wait for it to finish executing the critical section.
In this chapter, we will have a look at how one can develop concurrent
code in which processes do not have to wait for other processes to complete their
ongoing operations.

@@

There are also various algorithms where read-only access is wait-free
but updates do require a lock---these are useful in situations where
most accesses are read-only and the performance of updates is less
critical. %~\cite{}.
\autoref{fig:lst1} and \autoref{fig:lst2} implements an ordered linked
list of integers without duplicates.
There are two update operations:
values can be added using \texttt{lst\_insert} or deleted using
\texttt{lst\_remove}.
There is also a read-only operation: \texttt{lst\_contains} checks if
a particular value is in the list.
\glsadd{linearizable}%

@@

\begin{figure}
\begin{code}
\harmonysource{lst1}
\end{code}
\caption{\harmonylink{code/lst1.hny} List with non-blocking read operation, part 1.}
\label{fig:lst1}
\end{figure}

\begin{figure}
\begin{code}
\harmonysource{lst2}
\end{code}
\caption{\harmonylink{code/lst2.hny} List with non-blocking read operation, part 2.}
\label{fig:lst2}
\end{figure}
